---
title: "Food Hygiene data in OpenStreetMap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(colorspace)
library(lubridate)
library(scales)
library(targets)
library(tidyverse)

source("R/functions.R")

fhrs_summary_by_date <- tar_read(stats_by_district) %>%
  select(-stats_osm) %>%
  unnest(stats_fhrs) %>%
  group_by(date, old, statistic) %>%
  summarise(value = sum(value), .groups = "drop")

new_data_start <- fhrs_summary_by_date %>% filter(!old) %>% pull(date) %>% min()
latest_date <- fhrs_summary_by_date %>% pull(date) %>% max()

latest_total_establishments <- fhrs_summary_by_date %>%
  filter(date == max(date)) %>%
  summarise(value = sum(value)) %>%
  pull(value)

latest_matched_establishments <- fhrs_summary_by_date %>%
  filter(
    date == max(date),
    statistic == "matched_same_postcodes"
  ) %>%
  pull(value)

latest_fhrs_matched_percentage <- round(
  (latest_matched_establishments / latest_total_establishments) * 100,
  digits = 1
)

latest_fhrs_stats_by_district <- tar_read(stats_by_district) %>%
  select(-stats_osm) %>%
  unnest(stats_fhrs) %>%
  filter(date == max(date))
  # left_join(tar_read(ons_districts), by = c("district_code" = "code"))

latest_osm_stats_by_district <- tar_read(stats_by_district) %>%
  select(-stats_fhrs) %>%
  unnest(stats_osm) %>%
  filter(date == max(date))

theme_update(
  plot.background = element_rect(colour = "black"),
  plot.margin = unit(rep(0.75, 4), "lines")
)
```

# Matching progress

The graph below shows the progress made in matching Food Hygiene Rating Scheme establishments to OpenStreetMap objects using the `fhrs:id` and `addr:postcode` or `not:addr:postcode` tags. Matching objects are considered to have the same postcodes when the FHRS postcode matches either the OpenStreetMap object's `addr:postcode` or its `not:addr:postcode`.

```{r}
fhrs_statistic_defs <- get_fhrs_statistic_defs()

fhrs_summary_by_date %>%
  left_join(fhrs_statistic_defs, by = "statistic") %>%
  ggplot(aes(date, value, fill = label)) +
  geom_area() +
  scale_x_date(
    date_breaks = "1 year",
    date_minor_breaks = "1 month",
    labels = year
  ) +
  scale_y_continuous(labels = label_comma()) +
  scale_fill_manual(
    values = fhrs_statistic_defs %>% select(label, colour) %>% deframe()
  ) +
  guides(fill = guide_legend(reverse = TRUE, nrow = 2)) +
  labs(
    title = "Matching FHRS establishments with OpenStreetMap objects",
    x = "Date",
    y = "Number of FHRS establishments",
    fill = NULL,
    caption = paste(
      "Establishments in Northern Ireland and unmatched establishments with no",
      "location not included prior to", strftime(new_data_start, '%e %b %Y')
    )
  ) +
  theme(legend.position = "bottom")
```

**`r latest_fhrs_matched_percentage`%** of FHRS establishments (`r label_comma()(latest_matched_establishments)` out of `r label_comma()(latest_total_establishments)`) have been successfully matched to OpenStreetMap objects as of `r strftime(latest_date, '%A %e %B %Y')`.

```{r}
matched_prop_by_district <- latest_fhrs_stats_by_district %>%
  # ignore unmatched FHRS establishments without a location
  filter(statistic != "unmatched_without_location") %>%
  group_by(district_code) %>%
  mutate(total_establishments = sum(value)) %>%
  ungroup() %>%
  filter(statistic == "matched_same_postcodes") %>%
  mutate(prop_matched_same_postcodes = value / total_establishments)

matched_prop_by_district %>%
  ggplot(aes(prop_matched_same_postcodes)) +
  geom_jitter(aes(y = 0.5), height = 0.5) +
  geom_boxplot(aes(y = -1), coef = 999, outlier.shape = NA) +
  scale_x_continuous(labels = percent_format(), breaks = seq(0, 1, 0.2)) +
  scale_y_continuous(limits = c(-2, 1.5)) +
  labs(
    title = "Percentage of FHRS establishments matched in each district",
    subtitle = "Each dot represents one local authority district",
    x = "Percentage of FHRS establishments matched with same postcodes",
    y = NULL,
    caption = "Ignoring unmatched FHRS establishments without a location"
  ) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

The median percentage of FHRS establishments matched in a district is `r median(matched_prop_by_district$prop_matched_same_postcodes) %>% percent_format()()`. A quarter of districts have `r quantile(matched_prop_by_district$prop_matched_same_postcodes, 0.25) %>% percent_format()()` or less of their FHRS establishments matched but `r matched_prop_by_district %>% filter(prop_matched_same_postcodes >= 0.9) %>% nrow()` out of `r nrow(matched_prop_by_district)` districts have at least 90% of their FHRS establishments matched. (These statistics ignore unmatched FHRS establishments without a location and only consider matches where the FHRS establishment's postcode is the same as the OpenStreetMap object's `addr:postcode` or `not:addr:postcode`.)

# Matching issues

```{r}
fhrs_id_mismatches_by_district <- latest_osm_stats_by_district %>%
  filter(statistic == "mismatched") %>%
  slice_max(value, n = 20, with_ties = FALSE) %>%
  left_join(tar_read(ons_districts), by = c("district_code" = "code")) %>%
  mutate(name_ons = fct_reorder(name_ons, value))

fhrs_id_mismatches_by_district %>%
  ggplot(aes(value, name_ons)) +
  geom_col(fill = "#F15854") +
  labs(
    title = "OpenStreetMap objects with incorrect fhrs:id tags",
    subtitle = "by local authority district",
    x = "Number of OpenStreetMap objects",
    y = NULL
  )
```

`r fhrs_id_mismatches_by_district %>% slice_max(value, with_ties = FALSE) %>% pull(name_ons)` has `r fhrs_id_mismatches_by_district %>% slice_max(value, with_ties = FALSE) %>% pull(value)` OpenStreetMap objects with incorrect `fhrs:id` tags as of `r strftime(latest_date, '%A %e %B %Y')`.

```{r}
fhrs_matches_different_postcodes_by_district <- latest_osm_stats_by_district %>%
  filter(statistic == "matched_different_postcodes") %>%
  slice_max(value, n = 20, with_ties = FALSE) %>%
  left_join(tar_read(ons_districts), by = c("district_code" = "code")) %>%
  mutate(name_ons = fct_reorder(name_ons, value))

fhrs_matches_different_postcodes_by_district %>%
  ggplot(aes(value, name_ons)) +
  geom_col(fill = "#F15854") +
  labs(
    title = "Matched objects with different postcodes",
    subtitle = "by local authority district",
    x = "Number of OpenStreetMap objects",
    y = NULL
  )
```

`r fhrs_matches_different_postcodes_by_district %>% slice_max(value, with_ties = FALSE) %>% pull(name_ons)` has `r fhrs_matches_different_postcodes_by_district %>% slice_max(value, with_ties = FALSE) %>% pull(value)` matched OpenStreetMap objects with a different postcode from their corresponding FHRS establishment as of `r strftime(latest_date, '%A %e %B %Y')`. Postcode errors in the FHRS data can be noted using the `not:addr:postcode` tag on the relevant OpenStreetMap object.
